\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}

### 2023-01-03 submission/R1/manuscript.tex
minor formatting 2

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4084}
\usepackage{lineno}

@@

Participants were first given a brief demographic survey, which collected information regarding their \textbf{age, gender, country of birth, ethnicity and highest attained education level}. This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block. Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{24 items, 34} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{25 items, 35} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism). Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials). In total, each participant underwent 1340 trials of which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key (left/right, or up/down depending on the illusion type). For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest. \textbf{All trials were required to be completed within a single-session (total experiment duration: \textasciitilde55 minutes)}. The task was implemented using \emph{jsPsych}\textsuperscript{36} and was \textbf{hosted on \emph{Pavlovia} } \href{https://pavlovia.org/}{(\textbf{https://pavlovia.org/})}. The set of instructions for each illusion type is available in the experiment code.

@@

The analysis was carried out using \emph{R 4.2}\textsuperscript{40}, \emph{brms}\textsuperscript{41}, the \emph{tidyverse}\textsuperscript{42}, and the \emph{easystats} collection of packages\textsuperscript{43--46}. As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results\textsuperscript{based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}, see 47}.

@@

The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)). We fitted Bayesian linear models with the education level entered as a monotonic predictor\textsuperscript{appropriate for ordinal variables, 49}, which yielded no significant effects. For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model. These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions (\textbf{Figure 4}).

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4085}
\usepackage{lineno}

@@

Participants were first given a brief demographic survey, which collected information regarding their \textbf{age, gender, country of birth, ethnicity and highest attained education level}. This was followed by a practice series of illusions, after which the first series of 10 illusion blocks was presented in a randomized order, with a further randomization of the stimuli order within each block. Following this first series of blocks, two personality questionnaires were administered, the \emph{IPIP6}\textsuperscript{34} - measuring 6 ``normal'' personality traits (Extraversion, Openness, Conscientiousness, Agreeableness, Neuroticism and Honesty-Humility), and the \emph{PID-5}\textsuperscript{35} - measuring 5 ``pathological'' personality traits (Disinhibition, Antagonism, Detachment, Negative Affect and Psychoticism). Next, the second series of 10 illusion blocks was presented (with new randomized orders of blocks and trials). In total, each participant underwent 1340 trials of which they had to respond ``as fast as possible without making errors'' (i.e., an explicit double constraint to mitigate the inter-individual variability in the speed-accuracy trade off) by pressing the correct arrow key (left/right, or up/down depending on the illusion type). For instance, in the Müller-Lyer block, participants had to answer which one of the upper or bottom target line was the longest. \textbf{All trials were required to be completed within a single-session (total experiment duration: \textasciitilde55 minutes)}. The task was implemented using \emph{jsPsych}\textsuperscript{36} and was \textbf{hosted on \emph{Pavlovia} } \href{https://pavlovia.org/}{(\textbf{https://pavlovia.org/})}. The set of instructions for each illusion type is available in the experiment code.

@@

The analysis was carried out using \emph{R 4.2}\textsuperscript{40}, \emph{brms}\textsuperscript{41}, the \emph{tidyverse}\textsuperscript{42}, and the \emph{easystats} collection of packages\textsuperscript{43--46}. As all the full results have been made available (see \textbf{Data Availability}), we will focus here on the significant results
(based on the Bayes Factor \emph{BF} or the Probability of Direction \emph{pd}\textsuperscript{47}).

@@

The Bayesian t-tests on the effect of sex suggested anecdotal to moderate evidence in favour of the null effect for all scores, with the exception of the sensitivity to the Zöllner illusion, which was higher in males as compared to females (\(\Delta=-0.37\), 95\% CI {[}-0.62, -0.13{]}, \(BF_{10} = 12.74\)). We fitted Bayesian linear models with the education level entered as a monotonic predictor
(appropriate for ordinal variables\textsuperscript{49}{]})
which yielded no significant effects. For age, we fitted two types of models for each score, one general additive models (GAM) and a 2nd order polynomial model. These consistently suggested a significant positive linear relationship between age and Factor \emph{i} (\(pd=100\%\)), as well as the sensitivity to Müller-Lyer (\(pd=100\%\)), Vertical-Horizontal (\(pd=100\%\)), Zöllner (\(pd=100\%\)) and Ebbinghaus (\(pd=99\%\)) illusions (\textbf{Figure 4}).
